---
title: "Mexico FIP Risk analysis"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(fipriskanalysis)

#system("sudo apt-get install -y fftw3", wait = TRUE)
#system("sudo apt-get remove libfftw3-dev", wait = TRUE)
#BiocManager::install("EBImage")

```

Get life history data from Google sheet. This code will get the corresponding sheet and save it as *.rda in data folder

```{r get fip data, include=FALSE}

sheetlink = "https://docs.google.com/spreadsheets/d/1HoWV3HponuSLKvF1UyBOFXD-gbdwtiuhdui7OUnybZ4/edit#gid=2103071940"
# 
fipinventory <- read_googlesheet(sheetlink, sheetname = "inventario_fips_preeval", coltypes = "ccccccddDciccc")
usethis::use_data(fipinventory, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(fipinventory)
cat(data.description, file=here::here("R",paste0("data-","fipinventory.R")))

localities <- read_googlesheet(sheetlink, sheetname = "localidades_pno", coltypes = NULL)
#localities <- read_googlesheet(sheetlink, sheetname = "localidades_preeval")
usethis::use_data(localities, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(localities)
cat(data.description, file=here::here("R",paste0("data-","localities.R")))

lifehistorytraits <- read_googlesheet(sheetlink, sheetname = "life_history_traits", coltypes = NULL)
usethis::use_data(lifehistorytraits, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(lifehistorytraits)
cat(data.description, file=here::here("R",paste0("data-","lifehistorytraits.R")))

lifehistorydata <- read_googlesheet(sheetlink, sheetname = "datos_caracteristicas", coltypes = NULL)
usethis::use_data(lifehistorydata, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(lifehistorydata)
cat(data.description, file=here::here("R",paste0("data-","lifehistorydata.R")))

#updates package documentation
devtools::document()

#rebuild and load the package
devtools::load_all() # restarts and loads

 
```


Create data tables

```{r}

data("comunidadescosteras")
data("localities")

fiplocal.bc.id <- localities %>%
    dplyr::filter(fip_id_number %in% c(9158, 12856, 12947, 8040)) %>%
    dplyr::select(fip_id_number, COM_ID, locality, municipality, state) %>%
    dplyr::filter(COM_ID!="NA") %>% 
    dplyr::distinct(COM_ID) %>% 
    dplyr::pull(COM_ID)
  
comunidades.edos.bc <- estadoscosteros %>% 
  dplyr::filter(COM_ID %in% fiplocal.bc.id) %>% 
  dplyr::select(COM_ID, NOM_ENT, NOM_MUN, NOM_LOC, lon_deg, lon_min, lon_sec, lat_deg, lat_min, lat_sec) 

readr::write_csv(comunidades.edos.bc, here::here("outputs","Tabla_datos_posicion_geografica_comunidades.csv"))


```


Estimate vulnerability, for its different components
```{r get vulnerability}

#get data community vulnerability
data("adaptivecapacitydata")
data("localities")

fiplocal.bc.id <- localities %>%
    dplyr::filter(fip_id_number %in% c(9158, 12856, 12947, 8040)) %>%
    dplyr::select(fip_id_number, COM_ID, locality, municipality, state) %>%
    dplyr::filter(COM_ID!="NA") %>% 
    dplyr::distinct(COM_ID) %>% 
    dplyr::pull(COM_ID)

#composicion de la poblacion
comppob <- c("prop_hogares_mujeres","prop_ocupada", "prop_educacion", "grado_escolar","prop_menor")
pobreza <- c("prop_sin_escolar", "pob_sin_salud","prop_hog_sbienes","prop_desocup","irs","prop_alim")
hogares <- c("hogares_ptierra", "hogares_recamaras", "hogares_servicios", "hogares_eelectrica", "hogares_drenaje")
infraestructura <- c("tot_loc_subsidio", "hogares_aguaent", "long_cam" ,"disp_agua", "dist_via", "distancia")
pobindigena <- c("prop_pind", "prop_inmig", "prop_inana", "prop_indes")
disrupcion <- c("cam_pe", "cam_mig", "cam_pobsol", "cam_hogfem", "cam_religion")

subsidios.dat <- readr::read_csv(here::here("data-raw","BIENPESCA.csv")) %>% 
  dplyr::mutate(tot_subsidio = (`RECURSO FEDERAL APORTADO`+`RECURSO ESTATAL APORTADO`)) %>% 
  dplyr::group_by(COM_ID) %>% 
  dplyr::summarise(tot_loc_subsidio = sum(tot_subsidio)) %>% 
  dplyr::filter(COM_ID %in% fiplocal.bc.id) 

adaptive.bc <- adaptivecapacitydata %>% 
  dplyr::filter(COM_ID %in% fiplocal.bc.id) %>% 
  dplyr::left_join(subsidios.dat, by="COM_ID") %>% 
  dplyr::mutate(tot_loc_subsidio = dplyr::if_else(is.na(tot_loc_subsidio),0,tot_loc_subsidio))

comppob.pca <- pca_factor(thisdata = adaptive.bc, datasetname = "composicion_poblacion", namesdata = comppob) 
pobreza.pca <- pca_factor(thisdata = adaptive.bc, datasetname = "pobreza", namesdata = pobreza)
hogares.pca <- pca_factor(thisdata = adaptive.bc, datasetname = "hogares", namesdata = hogares)
infraestructura.pca <- pca_factor(thisdata = adaptive.bc, datasetname = "infraestructura", namesdata = infraestructura)
pobindigena.pca <- pca_factor(thisdata = adaptive.bc, datasetname = "poblacion_indigena", namesdata = pobindigena)
disrupcion.pca <- pca_factor(thisdata = adaptive.bc, datasetname = "disrupcion", namesdata = disrupcion)

vulnerabilidad.pca <- dplyr::bind_rows(comppob.pca, pobreza.pca, hogares.pca, infraestructura.pca, pobindigena.pca, disrupcion.pca)

readr::write_csv(vulnerabilidad.pca,here::here("outputs","vulnerability_pca_results.csv"))  

all.indicators <- c(comppob, pobreza, hogares, infraestructura, pobindigena, disrupcion) 

vul.data.set <- adaptive.bc %>%
    dplyr::select(any_of(c("NOM_ENT", "NOM_MUN","NOM_LOC","dec_lon","dec_lat", all.indicators))) 

readr::write_csv(vul.data.set, here::here("outputs","Tabla2_a_vulnerabilidad_indicadores.csv"))

vul.data.set.in <- adaptive.bc %>%
  dplyr::select(COM_ID, any_of(all.indicators)) %>% 
  dplyr::mutate(across(prop_hogares_mujeres:prop_alim, ~ 1 - .x)) %>% 
  dplyr::mutate(across(hogares_recamaras:hogares_drenaje, ~ 1 - .x)) %>%   
  dplyr::mutate(across(dist_via:distancia, ~.x * -1)) %>% 
  dplyr::mutate(across(tot_loc_subsidio:disp_agua, ~ 1 - .x)) %>% 
  dplyr::mutate(across(cam_pe:cam_religion, ~ 1 - .x)) 

loc.fip.vul <- localities %>%
  dplyr::filter(fip_id_number %in% c(9158, 12856, 12947, 8040)) %>% 
  dplyr::select(COM_ID, fip_name, fip_id_number) %>% 
  dplyr::left_join(vul.data.set.in, by=c("COM_ID")) %>% 
  dplyr::filter(!is.na(cam_hogfem)) %>% 
  tidyr::pivot_longer(cols=4:ncol(vul.data.set.in), names_to="indicator",values_to="value") %>% 
  dplyr::group_by(fip_name, fip_id_number, indicator) %>% 
  dplyr::summarise(fip_value = mean(value)) %>% 
  tidyr::pivot_wider(names_from = indicator, values_from = fip_value) %>% 
  dplyr::ungroup()

process.vul <- loc.fip.vul %>% 
  dplyr::select(dplyr::where(is.numeric)) %>%
  dplyr::select(-fip_id_number) %>% 
  caret::preProcess(., method=c("range"))

vul.data.set.sc <- loc.fip.vul %>% 
  dplyr::select(dplyr::where(is.numeric)) %>%
  dplyr::select(-fip_id_number) %>% 
  stats::predict(process.vul, .) %>% 
  dplyr::bind_cols(fip_name=loc.fip.vul$fip_name, fip_id_number=loc.fip.vul$fip_id_number, .)

scaled.vul <- scale_var(vul.data.set.sc, NA)
         
  
readr::write_csv(scaled.vul, here::here("outputs","scaled_vulnerability.csv"))


```

Estimate exposure
```{r get exposure}
data("adaptivecapacitydata")
data("sensitivitydata")
data("localities")
data("fipinventory")

fiplocal.bc.id <- localities %>%
    dplyr::filter(fip_id_number %in% c(9158, 12856, 12947, 8040)) %>%
    dplyr::select(fip_id_number, COM_ID, locality, municipality, state) %>%
    dplyr::filter(COM_ID!="NA") %>% 
    dplyr::distinct(COM_ID) %>% 
    dplyr::pull(COM_ID)

fip.catch <- fipinventory %>% 
  dplyr::filter(fip_id_number %in% c(9158, 12856, 12947, 8040)) %>%
  dplyr::select(fip_id_number, species, estimated_total_fip_landings_mt, estimated_total_fishery_landings_mt) %>% 
  dplyr::left_join(localities, by="fip_id_number", multiple = "all") %>% 
  dplyr::group_by(COM_ID) %>% 
  dplyr::summarise(tot_estimated_fip_landings_mt = mean(estimated_total_fip_landings_mt) , tot_estimated_fishery_landings_mt= mean(estimated_total_fishery_landings_mt))

#composicion de la poblacion
pesqueria <- c("prop_pesca","neg_pesca_per_capita","tot_estimated_fip_landings_mt","tot_estimated_fishery_landings_mt","tot_captura_loc_per_capita")
empleo <- c("pob_ecactiva", "pob_empleada","grad_esc","in_compeco","in_marg","pob_salud")

sensitivity.bc <- sensitivitydata %>% 
  dplyr::filter(COM_ID %in% fiplocal.bc.id) %>% 
  dplyr::left_join(fip.catch, by="COM_ID")

adaptive.bc <- adaptivecapacitydata %>% 
   dplyr::filter(COM_ID %in% fiplocal.bc.id)
  
pesqueria.pca <- pca_factor(thisdata = sensitivity.bc, datasetname = "pesqueria", namesdata = pesqueria) 
empleo.pca <- pca_factor(thisdata = adaptive.bc, datasetname = "empleo", namesdata = empleo)

exposicion.pca <- dplyr::bind_rows(pesqueria.pca, empleo.pca)

readr::write_csv(exposicion.pca,here::here("outputs","exposure_pca_results.csv"))  

data.set.em <- adaptive.bc %>%
  dplyr::mutate(NOM_MUN = tolower(stringi::stri_trans_general(str = NOM_MUN, id = "Latin-ASCII"))) %>% # quitar los acentos
  dplyr::mutate(NOM_ENT = tolower(stringi::stri_trans_general(str = NOM_ENT, id = "Latin-ASCII"))) %>% 
  dplyr::mutate(NOM_LOC = tolower(stringi::stri_trans_general(str = NOM_LOC, id = "Latin-ASCII"))) %>%  
  dplyr::select(any_of(c("NOM_ENT", "NOM_MUN","NOM_LOC","COM_ID","dec_lon","dec_lat",empleo)))

data.set.pes <- sensitivity.bc %>%
    dplyr::select(any_of(c("NOM_ENT", "NOM_MUN","NOM_LOC","COM_ID","dec_lon","dec_lat",pesqueria)))

data.set.exp <- data.set.em %>% 
  dplyr::left_join(data.set.pes, by=c("NOM_ENT", "NOM_MUN","NOM_LOC", "COM_ID"))

readr::write_csv(data.set.exp, here::here("outputs","Tabla2_b_exposicion_indicadores.csv"))

exp.data.set.in <- data.set.exp %>%
    dplyr::select(-NOM_ENT, -NOM_MUN, -NOM_LOC, -dec_lon, -dec_lat) 

fip.catch.id <- fipinventory %>% 
  dplyr::filter(fip_id_number %in% c(9158, 12856, 12947, 8040)) %>%
  dplyr::distinct(fip_id_number, estimated_total_fip_landings_mt, estimated_total_fishery_landings_mt)

loc.fip.exp <- localities %>%
  dplyr::filter(fip_id_number %in% c(9158, 12856, 12947, 8040)) %>% 
  dplyr::select(COM_ID, fip_name, fip_id_number) %>% 
  dplyr::left_join(exp.data.set.in, by=c("COM_ID")) %>% 
  dplyr::filter(!is.na(pob_ecactiva)) %>% 
  dplyr::select(-tot_estimated_fip_landings_mt, -tot_estimated_fishery_landings_mt) %>% 
  tidyr::pivot_longer(cols=pob_ecactiva:tot_captura_loc_per_capita, names_to="indicator",values_to="value") %>% 
  dplyr::group_by(fip_name, fip_id_number, indicator) %>% 
  dplyr::summarise(fip_value = mean(value)) %>% 
  tidyr::pivot_wider(names_from = indicator, values_from = fip_value) %>% 
  dplyr::ungroup() %>% 
  dplyr::left_join(fip.catch.id, by=c("fip_id_number"))

process.exp <- loc.fip.exp %>% 
  dplyr::select(dplyr::where(is.numeric)) %>%
  dplyr::select(-fip_id_number) %>% 
  caret::preProcess(., method=c("range"))

exp.data.set.sc <- loc.fip.exp %>% 
  dplyr::select(dplyr::where(is.numeric)) %>%
  dplyr::select(-fip_id_number) %>% 
  stats::predict(process.exp, .) %>% 
  dplyr::bind_cols(fip_name=loc.fip.exp$fip_name, fip_id_number=loc.fip.exp$fip_id_number, .)

scaled.exp <- scale_var(exp.data.set.sc, NA)

readr::write_csv(scaled.exp, here::here("outputs","scaled_exposure.csv"))


#factorscores.impact is saved in pca_factor() and has the PC loadings by community

```


Obtain life history data by species

```{r}

#FIP names and numbers
#Mexico Bahia de Los Angeles octopus - trap/diver-caught/hand gathered	9158
#Mexico Baja California red rock lobster - trap	12856
#Mexico Baja California red sea urchin - diver-caught	12947
#Mexico North Pacific barred sand bass - pot/trap 8040

data("fipinventory")
data("lifehistorydata")
data("localities")
data("lifehistorytraits")

lifehistory.trait <- get_lifehistory(lifehistorydata, localities, fipinventory, lifehistorytraits)

readr::write_csv(lifehistory.trait, here::here("outputs","lifehistory_traits.csv"))


```



Estimate statistics from changes in distribution by species and future scenario
```{r}

#estimate changes in distribution

future.change <- c("ssp126", "ssp245", "ssp585")

species <- c("Mesocentrotus_franciscanus", "Bimaculatus","Octopus_hubbsorum","Paralabrax_nebulifer","Panulirus_interruptus")
dirnames <- c("erizo","Pulpos","Pulpos","verdillo","Langosta")

#species <- c("Mesocentrotus_franciscanus", "Octopus_hubbsorum","Paralabrax_nebulifer","Panulirus_interruptus")
#dirnames <- c("erizo","Pulpos", "verdillo","Langosta")

dist.data <- list()

for(eachspecies in 1:length(species)) {
  
  eachspeciesname <- species[eachspecies]
  eachdirname <- dirnames[eachspecies]
  print(eachspeciesname)
  print(eachdirname)
  raster.present <- terra::rast(here::here("data-raw","modelos", eachdirname,paste0(eachspeciesname,"_avg.asc")))
  crs.eqap <- enmSdmX::getCRS('EE Americas') # Equal area projection Americas, use getCRS() to see all projections available
  raster.present.proj <- terra::project(raster.present, crs.eqap)

  dist.stats <- lapply(future.change, calc_distchange, raster.present.proj, eachspeciesname, crs.eqap, eachdirname) 
  
  dist.data[[eachspecies]] <- dist.stats
 }

  
dist.table <- dist.data %>% 
  dplyr::bind_rows() %>% 
  tidyr::pivot_longer(cols=centroid_velocity:rankCor, names_to = "indice") %>% 
  dplyr::mutate(species=gsub("_"," ", species),
                species= dplyr::if_else(species == "Bimaculatus", "Octopus bimaculatus", species))
  

readr::write_csv(dist.table, here::here("outputs","distribution_table.csv"))

```

Estimate species sensitivity

```{r get sensitivity}


lifehistory.trait <- readr::read_csv(here::here("outputs","lifehistory_traits.csv"))

dist.table <- readr::read_csv(here::here("outputs","distribution_table.csv")) %>% 
  tidyr::pivot_wider(names_from = indice, values_from = "value")

#composicion de la poblacion
lifehistory <- lifehistory.trait %>% 
  dplyr::select(-species) %>% 
  names()
  
distrib <- dist.table %>% 
  dplyr::select(-species, -scenario) %>% 
  names()

scaled.ssp126 <- scale_var(dist.table, "ssp126")
scaled.ssp245 <- scale_var(dist.table, "ssp245")
scaled.ssp585 <- scale_var(dist.table, "ssp585")

scaled.lifehistory <- scale_var(datatable=lifehistory.trait, sspscenario=NA)

scaled.distribution <- dplyr::bind_rows(scaled.ssp126, scaled.ssp245, scaled.ssp585)

readr::write_csv(scaled.distribution, here::here("outputs","Tabla2_c_sensibilidad_distribucion.csv"))  

readr::write_csv(scaled.lifehistory, here::here("outputs","Tabla2_d_sensibilidad_historiavida.csv"))

readr::write_csv(scaled.distribution, here::here("outputs","scaled_distribution.csv"))  

readr::write_csv(scaled.lifehistory, here::here("outputs","scaled_lifehistory.csv"))

#factorscores.impact is saved in pca_factor() and has the PC loadings by community

```



Estimate hazard
```{r get hazard}

data("exposurehistorical")
data("exposuressp126")
data("exposuressp245")
data("exposuressp585")
data("localities")

#exposure by communities
peligroterrestre <- c("impacto_antropogenico","impacto_humano","temperatura_promedio_anual","intervalo_diurno", "precipitacion_promedio","riesgo_inundaciones")
peligromarino <- c("riesgo_huracanes","marea_roja", "temperatura_superficial_del_mar", "productividad_primaria","oxigeno_disuelto")

fiplocal.bc.id <- localities %>%
    dplyr::filter(fip_id_number %in% c(9158, 12856, 12947, 8040)) %>%
    dplyr::select(fip_id_number, COM_ID, locality, municipality, state) %>%
    dplyr::filter(COM_ID!="NA") %>% 
    dplyr::distinct(COM_ID) %>% 
    dplyr::pull(COM_ID)

expohist.bc <- exposurehistorical %>% 
  dplyr::filter(COM_ID %in% fiplocal.bc.id) %>% 
  dplyr::mutate(escenario = "historico") %>% 
  #tranpose the directionaly so that it is correct when standarized
  dplyr::mutate(impacto_antropogenico=impacto_antropogenico * -1)

expossp126.bc <- exposuressp126 %>% 
  dplyr::filter(COM_ID %in% fiplocal.bc.id)%>% 
  dplyr::mutate(escenario = "ssp126") %>% 
  #tranpose the directionaly so that it is correct when standarized
  dplyr::mutate(impacto_antropogenico=impacto_antropogenico * -1)

expossp245.bc <- exposuressp245 %>% 
  dplyr::filter(COM_ID %in% fiplocal.bc.id)%>% 
  dplyr::mutate(escenario = "ssp245") %>% 
  #tranpose the directionaly so that it is correct when standarized
  dplyr::mutate(impacto_antropogenico=impacto_antropogenico * -1)

expossp585.bc <- exposuressp585 %>% 
  dplyr::filter(COM_ID %in% fiplocal.bc.id)%>% 
  dplyr::mutate(escenario = "ssp585") %>% 
  #tranpose the directionaly so that it is correct when standarized
  dplyr::mutate(impacto_antropogenico=impacto_antropogenico * -1)

#Continue here, for some reason it is not saving the row names in the pca contribution

peligrote.hist.pca <- pca_factor(thisdata = expohist.bc, datasetname = "peligro terrestre_hist", namesdata = peligroterrestre) 
peligroma.hist.pca <- pca_factor(thisdata = expohist.bc, datasetname = "peligro marino_hist", namesdata = peligromarino) 

peligrote.126.pca <- pca_factor(thisdata = expossp126.bc, datasetname = "peligro terrestre_ssp126", namesdata = peligroterrestre) 
peligroma.126.pca <- pca_factor(thisdata = expossp126.bc, datasetname = "peligro marino_ssp126", namesdata = peligromarino) 

peligrote.245.pca <- pca_factor(thisdata = expossp245.bc, datasetname = "peligro terrestre_ssp245", namesdata = peligroterrestre) 
peligroma.245.pca <- pca_factor(thisdata = expossp245.bc, datasetname = "peligro marino_ssp245", namesdata = peligromarino) 

peligrote.585.pca <- pca_factor(thisdata = expossp585.bc, datasetname = "peligro terrestre_ssp585", namesdata = peligroterrestre) 
peligroma.585.pca <- pca_factor(thisdata = expossp585.bc, datasetname = "peligro marino_ssp585", namesdata = peligromarino) 

hazard.pca.coms <- dplyr::bind_rows(peligrote.hist.pca, peligroma.hist.pca, peligrote.126.pca, peligroma.126.pca, 
                                   peligrote.245.pca, peligroma.245.pca, peligrote.585.pca, peligroma.585.pca) %>% 
                  tidyr::separate_wider_delim(cols=data_name, delim="_", names=c("data_name","scenario"))
  

readr::write_csv(hazard.pca.coms,here::here("outputs","hazard_pca_results_coms.csv"))  

exp.scenarios <- c(expohist.bc, expossp126.bc, expossp245.bc, expossp585.bc)
num.exp <- length(exp.scenarios)
names.scenarios <- c("historico", "ssp126", "ssp245", "ssp585")
id.cols <- expohist.bc %>% 
    dplyr::select(COM_ID, NOM_ENT, NOM_MUN, NOM_LOC)

expo.hist.mar <- select_var_tab(namesdata=peligromarino, datatable = expohist.bc, scenario = "historico", id.cols)
expo.ssp126.mar <- select_var_tab(namesdata=peligromarino, datatable = expossp126.bc, scenario = "ssp126", id.cols)
expo.ssp245.mar <- select_var_tab(namesdata=peligromarino, datatable = expossp245.bc, scenario = "ssp245", id.cols)
expo.ssp585.mar <- select_var_tab(namesdata=peligromarino, datatable = expossp585.bc, scenario = "ssp585", id.cols)

expo.mar <- dplyr::bind_rows(expo.hist.mar, expo.ssp126.mar, expo.ssp245.mar, expo.ssp585.mar) 

readr::write_csv(expo.mar, here::here("outputs","Tabla2_e_peligro_marino.csv"))

expo.hist.terr <- select_var_tab(namesdata=peligroterrestre, datatable = expohist.bc, scenario = "historico", id.cols)
expo.ssp126.terr <- select_var_tab(namesdata=peligroterrestre, datatable = expossp126.bc, scenario = "ssp126", id.cols)
expo.ssp245.terr <- select_var_tab(namesdata=peligroterrestre, datatable = expossp245.bc, scenario = "ssp245", id.cols)
expo.ssp585.terr <- select_var_tab(namesdata=peligroterrestre, datatable = expossp585.bc, scenario = "ssp585", id.cols)

expo.terr <- dplyr::bind_rows(expo.hist.terr, expo.ssp126.terr, expo.ssp245.terr, expo.ssp585.terr) 

readr::write_csv(expo.terr, here::here("outputs","Tabla2_f_peligro_terrestre.csv"))

loc.fip.expo <- localities %>%
  dplyr::filter(fip_id_number %in% c(9158, 12856, 12947, 8040)) %>% 
  dplyr::select(COM_ID, fip_name, fip_id_number) 


expo.hist.mar.sc <- scale_indi(datatable = expo.hist.mar, loc.fip.expo)
expo.ssp126.mar.sc <- scale_indi(datatable = expo.ssp126.mar, loc.fip.expo)
expo.ssp245.mar.sc <- scale_indi(datatable = expo.ssp245.mar, loc.fip.expo)
expo.ssp585.mar.sc <- scale_indi(datatable = expo.ssp585.mar, loc.fip.expo)

expo.mar.sc <- dplyr::bind_rows(expo.hist.mar.sc, expo.ssp126.mar.sc, expo.ssp245.mar.sc, expo.ssp585.mar.sc) %>% 
    dplyr::rename(tot_score_mar = tot_score) %>% 
    tidyr::pivot_longer(cols=3:ncol(.), names_to="indicator")

expo.hist.terr.sc <- scale_indi(datatable = expo.hist.terr, loc.fip.expo)
expo.ssp126.terr.sc <- scale_indi(datatable = expo.ssp126.terr, loc.fip.expo)
expo.ssp245.terr.sc <- scale_indi(datatable = expo.ssp245.terr, loc.fip.expo)
expo.ssp585.terr.sc <- scale_indi(datatable = expo.ssp585.terr, loc.fip.expo)

expo.terr.sc <- dplyr::bind_rows(expo.hist.terr.sc, expo.ssp126.terr.sc, expo.ssp245.terr.sc, expo.ssp585.terr.sc) %>% 
   dplyr::rename(tot_score_terr = tot_score) %>% 
  tidyr::pivot_longer(cols=3:ncol(.), names_to="indicator")

hazard.scaled <- expo.mar.sc %>% 
  dplyr::bind_rows(expo.terr.sc) %>% 
  tidyr::pivot_wider(names_from = indicator, values_from = value)

readr::write_csv(hazard.scaled, here::here("outputs","scaled_hazard.csv"))

```


plot sensitivity 
```{r}

sens.dist <- readr::read_csv(here::here("outputs","scaled_distribution.csv"))  
sens.lifehistory <- readr::read_csv(here::here("outputs","scaled_lifehistory.csv"))  

#plot of species sensitivity

sens.data <- sens.lifehistory %>%
  dplyr::rename(tot_score_dist = tot_score) %>%
  dplyr::left_join(sens.dist, by = "species") %>%
  dplyr::mutate(
    image = dplyr::if_else(
      species == "Paralabrax nebulifer",
      "https://images.phylopic.org/images/eb6f54bf-3182-4d3a-9f16-602759c76db0/raster/512x253.png",
      dplyr::if_else(
        species == "Mesocentrotus franciscanus",
        "https://images.phylopic.org/images/0552ec39-a86b-4e23-b92a-7c1c38386e29/raster/1024x715.png",
        dplyr::if_else(
          species == "Octopus hubbsorum",
          "https://images.phylopic.org/images/f400b519-3564-4183-b4bd-c3b922cc7c5e/raster/512x373.png",
          dplyr::if_else(
            species == "Panulirus interruptus",
            "https://images.phylopic.org/images/d96e18f2-c9d6-4b2e-914c-678ed3c72a28/raster/759x1024.png",
            dplyr::if_else(
              species == "Octopus bimaculatus",
              "https://images.phylopic.org/images/85a4f59d-7a2e-4399-98e2-38d0edc75996/raster/512x348.png",
              "NA"
            )
          )
        )
      )
    )
  )

sp.cols <- c("#E69F00", "#56B4E9", "#873e23", "#F0E442", "#0072B2", "#999999")
  
sens.data.plot <- sens.data %>% 
  ggplot2::ggplot(ggplot2::aes(x=tot_score, y=tot_score_dist)) + 
  ggplot2::geom_point(ggplot2::aes(color = species),
    size = 2, 
    alpha = 0.8 # It's nice to add some transparency because there may be overlap.
  ) +
  ggplot2::scale_color_manual(values=sp.cols, name="Especie") +
  ggplot2::facet_wrap(.~ scenario, ncol = 1) +
  ggplot2::labs(
    title = "Sensibilidad relativa de especies en base a distribución \ne historia de vida",
    subtitle = "Valores mas altos indican mayor sensibilidad",
    x = "Sensibilidad relativa historia de vida",
    y = "Sensibilidad relativa distribución"
  ) +
  ggplot2::theme_linedraw() +
  ggplot2::scale_shape(guide="none") +
  ggimage::geom_image(ggplot2::aes(image=image), size=0.12, nudge_x = 0.1)+
  ggplot2::xlim(0,1.2)+
  ggplot2::ylim(-0.2,1.2)

sens.data.plot

ggplot2::ggsave("sensibilidad_especies.png", sens.data.plot, path = here::here("outputs","figures"), dpi=400, width= 7, height=9,scale =1)



```

Plot vulnerability, hazard and exposure

```{r}

sens.lifehistory <- readr::read_csv(here::here("outputs","scaled_lifehistory.csv"))  %>% 
dplyr::mutate(fip_id_number = c(12947, 12856, 8040, 9158, 9158)) %>% 
  dplyr::group_by(fip_id_number) %>% 
  dplyr::summarise(tot_score_lf = mean(tot_score)) 

# dist.lifehistory <- readr::read_csv(here::here("outputs","scaled_distribution.csv"))  %>% 
# dplyr::mutate(fip_id_number = c(12947, 12856, 8040, 9158, 9158)) %>% 
#   dplyr::group_by(fip_id_number) %>% 
#   dplyr::summarise(tot_score_dis = mean(tot_score))

vul.sc.data <-
  readr::read_csv(here::here("outputs", "scaled_vulnerability.csv"))  %>%
  dplyr::rename(tot_score_vul = tot_score) %>% 
  dplyr::select(-fip_id_number) %>% 
  tidyr::pivot_longer(cols = 2:ncol(.), names_to = "indicator", values_to = "value")%>%
    dplyr::mutate(FIP = dplyr::if_else(fip_name == "Mexico Bahia de Los Angeles octopus - trap/diver-caught/hand gathered", "Pulpo",
                                       dplyr::if_else(fip_name =="Mexico Baja California red sea urchin - diver-caught","Erizo rojo",
                                                      dplyr::if_else(fip_name =="Mexico North Pacific barred sand bass - pot/trap","Verdillo", "Langosta")))) %>% 
  dplyr::mutate(componente = "vulnerabilidad")

vul.tabla <- vul.sc.data %>% 
  dplyr::select(-fip_name) %>% 
  dplyr::mutate(indicator = dplyr::if_else(indicator== "fip_name", "nombre_fip",
                               dplyr::if_else(indicator=="estimated_total_fip_landings_mt", "captura_fip", 
                                              dplyr::if_else(indicator=="estimated_total_fishery_landings_mt", "produccion_pesquera", 
                                                             dplyr::if_else(indicator=="tot_score_exp", "valor_total_expo",
                                                                            dplyr::if_else(indicator=="tot_lf_catch", "valor_total_historia", indicator)))))) %>% 
  dplyr::rename(indicador=indicator, valor = value)

readr::write_csv(vul.tabla, here::here("outputs","Tabla 3_a_vulnerabilidad.csv"))


expo.sc.data <-
  readr::read_csv(here::here("outputs", "scaled_exposure.csv"))  %>%
  dplyr::rename(tot_score_exp = tot_score) %>% 
  dplyr::left_join(sens.lifehistory, by = "fip_id_number") %>% 
  dplyr::mutate(tot_lf_catch = tot_score_lf * (estimated_total_fip_landings_mt/estimated_total_fishery_landings_mt)) %>% 
  dplyr::mutate(tot_lf_catch = dplyr::if_else(is.nan(tot_lf_catch),0, tot_lf_catch))

process.num <- expo.sc.data %>%
    dplyr::select(tot_lf_catch) %>%
    caret::preProcess(., method=c("range"))

num.scale <- expo.sc.data %>%
    dplyr::select(tot_lf_catch) %>%
    stats::predict(process.num, .) 

expo.sc.data.lf <- expo.sc.data %>%
  dplyr::select(-tot_score_lf, -tot_lf_catch, -fip_id_number) %>% 
    dplyr::bind_cols(num.scale) %>% 
  tidyr::pivot_longer(cols = 2:ncol(.), names_to = "indicator", values_to = "value") %>% 
  dplyr::mutate(FIP = dplyr::if_else(fip_name == "Mexico Bahia de Los Angeles octopus - trap/diver-caught/hand gathered", "Pulpo",
                                       dplyr::if_else(fip_name =="Mexico Baja California red sea urchin - diver-caught","Erizo rojo",
                                                      dplyr::if_else(fip_name =="Mexico North Pacific barred sand bass - pot/trap","Verdillo", "Langosta")))) %>% 
   dplyr::mutate(componente = "exposición")

expo.tabla <- expo.sc.data.lf %>% 
  dplyr::select(-fip_name) %>% 
  dplyr::mutate(indicator = dplyr::if_else(indicator== "fip_name", "nombre_fip",
                               dplyr::if_else(indicator=="estimated_total_fip_landings_mt", "captura_fip", 
                                              dplyr::if_else(indicator=="estimated_total_fishery_landings_mt", "produccion_pesquera", 
                                                             dplyr::if_else(indicator=="tot_score_exp", "valor_total_expo",
                                                                            dplyr::if_else(indicator=="tot_lf_catch", "valor_total_historia", indicator)))))) %>% 
  dplyr::rename(indicador=indicator, valor = value)

readr::write_csv(expo.tabla, here::here("outputs","Tabla 3_b_exposicion.csv"))


hazard.sc.data <-
  readr::read_csv(here::here("outputs", "scaled_hazard.csv")) %>% 
  tidyr::pivot_longer(cols = 3:ncol(.), names_to = "indicator", values_to = "value")%>% 
  dplyr::mutate(FIP = dplyr::if_else(fip_name == "Mexico Bahia de Los Angeles octopus - trap/diver-caught/hand gathered", "Pulpo",
                                       dplyr::if_else(fip_name =="Mexico Baja California red sea urchin - diver-caught","Erizo rojo",
                                                      dplyr::if_else(fip_name =="Mexico North Pacific barred sand bass - pot/trap","Verdillo", "Langosta")))) %>% 
   dplyr::mutate(componente = "peligro")


hazard.tabla <- hazard.sc.data %>% 
  dplyr::select(-fip_name) %>% 
  dplyr::mutate(indicator = dplyr::if_else(indicator== "fip_name", "nombre_fip",
                               dplyr::if_else(indicator=="estimated_total_fip_landings_mt", "captura_fip", 
                                              dplyr::if_else(indicator=="estimated_total_fishery_landings_mt", "produccion_pesquera", 
                                                             dplyr::if_else(indicator=="tot_score_exp", "valor_total_expo",
                                                                            dplyr::if_else(indicator=="tot_lf_catch", "valor_total_historia", indicator)))))) %>% 
  dplyr::rename(indicador=indicator, valor = value)

readr::write_csv(hazard.tabla, here::here("outputs","Tabla 3_c_peligro.csv"))


vul.exp.data <- dplyr::bind_rows(vul.sc.data, expo.sc.data.lf)

hazard.sc <- hazard.sc.data %>%
  dplyr::distinct(escenario) %>%
  dplyr::pull(escenario)

  this.data <- hazard.sc.data %>% 
  dplyr::filter(escenario==eachscenario) %>% 
  dplyr::select(-escenario) %>% 
  dplyr::bind_rows(vul.exp.data) %>% 
  dplyr::select(-fip_name) 
  
  sp.cols <-
  c("#154c79",
             "#e28743",
             "#76b5c5",
             "#873e23",
             "#0072B2",
             "#56B4E9",
             "#999999")
  
comp.plot <- this.data %>%
  ggplot2::ggplot(ggplot2::aes(x = componente, y = value)) +
  ggplot2::geom_point(ggplot2::aes(color = componente),
                      size = 2,
                      alpha = 0.8) + # It's nice to add some transparency because there may be overlap.)
                      ggplot2::scale_color_manual(values = sp.cols, name = "Componente") +
                        ggplot2::labs(
                          title = "Componentes del riesgo a las pesquerías",
                          subtitle = "Puntos son diferentes indicadores",
                          x = "Componente",
                          y = "Valor"
                        ) +
                        ggplot2::theme_linedraw() +
                        ggplot2::scale_shape(guide = "none") +
                        ggplot2::theme(legend.position="bottom") +
                        ggplot2::facet_wrap(.~FIP, ncol=2) 
                      

ggplot2::ggsave(
  "componentes_pesqueria.png",
  comp.plot,
  path = here::here("outputs", "figures"),
  dpi = 400,
  width = 9.5,
  height = 10,
  scale = 1
)


```



Plot risk
```{r plot risk}

sens.lifehistory <- readr::read_csv(here::here("outputs","scaled_lifehistory.csv"))  %>% 
dplyr::mutate(fip_id_number = c(12947, 12856, 8040, 9158, 9158)) %>% 
  dplyr::group_by(fip_id_number) %>% 
  dplyr::summarise(tot_score_lf = mean(tot_score))

# dist.lifehistory <- readr::read_csv(here::here("outputs","scaled_distribution.csv"))  %>% 
# dplyr::mutate(fip_id_number = c(12947, 12856, 8040, 9158, 9158)) %>% 
#   dplyr::group_by(fip_id_number) %>% 
#   dplyr::summarise(tot_score_dis = mean(tot_score))

vul.sc.data <-
  readr::read_csv(here::here("outputs", "scaled_vulnerability.csv"))  %>%
  dplyr::rename(tot_score_vul = tot_score, loc_subsidio = tot_loc_subsidio)

expo.sc.data <-
  readr::read_csv(here::here("outputs", "scaled_exposure.csv"))  %>%
  dplyr::rename(tot_score_exp = tot_score) %>% 
  dplyr::left_join(sens.lifehistory, by = "fip_id_number") %>% 
  dplyr::mutate(tot_lf_catch = tot_score_lf * (estimated_total_fip_landings_mt/estimated_total_fishery_landings_mt)) %>% 
  dplyr::mutate(tot_lf_catch = dplyr::if_else(is.nan(tot_lf_catch),0, tot_lf_catch))

process.num <- expo.sc.data %>%
    dplyr::select(tot_lf_catch) %>%
    caret::preProcess(., method=c("range"))

num.scale <- expo.sc.data %>%
    dplyr::select(tot_lf_catch) %>%
    stats::predict(process.num, .) 

expo.sc.data.lf <- expo.sc.data %>%
  dplyr::select(-tot_score_lf, -tot_lf_catch) %>% 
    dplyr::bind_cols(num.scale)

hazard.sc.data <-
  readr::read_csv(here::here("outputs", "scaled_hazard.csv"))

vul.expo.dat <- vul.sc.data %>%
  dplyr::left_join(expo.sc.data.lf, by = c("fip_name", "fip_id_number")) %>%
  dplyr::select(-fip_id_number)

hazard.sc <- hazard.sc.data %>%
  dplyr::distinct(escenario) %>%
  dplyr::pull(escenario)

topsis.res <-
  lapply(hazard.sc, calc_topsis, hazard.sc.data, vul.expo.dat, sp.cols) %>%
  dplyr::bind_rows() %>%
  dplyr::mutate(escenario = dplyr::if_else(escenario == "historico", "histórico", escenario))

sp.cols <-
  c("#E69F00",
             "#F0E442",
             "#873e23",
             "#0072B2",
             "#56B4E9",
             "#999999")
             
topsis.plot <- topsis.res %>%
  ggplot2::ggplot(ggplot2::aes(x = escenario, y = score)) +
  ggplot2::geom_point(ggplot2::aes(color = FIP),
                      size = 2,
                      alpha = 0.8) + # It's nice to add some transparency because there may be overlap.)
                      ggplot2::scale_color_manual(values = sp.cols, name = "FIP") +
                        ggplot2::labs(
                          title = "Riesgo a las pesquerías FIP debido al cambio climático",
                          subtitle = "Valores mas altos indican mayor riesgo",
                          x = "Escenario",
                          y = "Riesgo a la pesquería"
                        ) +
                        ggplot2::theme_linedraw() +
                        ggplot2::scale_shape(guide = "none") +
                        ggplot2::theme(legend.position="bottom")
                      
                      # https://stackoverflow.com/questions/60917778/how-do-i-plot-an-image-from-phylopic-in-top-right-corner-of-my-ggplot-graph-in-r
                      verdillo <-
                        png::readPNG(
                          RCurl::getURLContent(
                            "https://images.phylopic.org/images/eb6f54bf-3182-4d3a-9f16-602759c76db0/raster/512x253.png"
                          ),
                          native = T
                        )
                      verdillo.grob <- grid::rasterGrob(verdillo, interpolate = TRUE)
                      
                      erizo <-
                        png::readPNG(
                          RCurl::getURLContent(
                            "https://images.phylopic.org/images/0552ec39-a86b-4e23-b92a-7c1c38386e29/raster/1024x715.png"
                          ),
                          native = T
                        )
                      erizo.grob <- grid::rasterGrob(erizo, interpolate = TRUE)
                      
                      pulpo <-
                        png::readPNG(
                          RCurl::getURLContent(
                            "https://images.phylopic.org/images/f060bcc7-5725-46f7-8276-14553af3707f/raster/1452x1536.png"
                          ),
                          native = T
                        )
                      pulpo.grob <- grid::rasterGrob(pulpo, interpolate = TRUE)
                      
                      langosta <-
                        png::readPNG(
                          RCurl::getURLContent(
                            "https://images.phylopic.org/images/d96e18f2-c9d6-4b2e-914c-678ed3c72a28/raster/759x1024.png"
                          ),
                          native = T
                        )
                      langosta.grob <- grid::rasterGrob(langosta, interpolate = TRUE)
                      
                      
topsis.plot.icons <- topsis.plot +
                        ggplot2::annotation_custom(
                          erizo.grob,
                          xmin = 3.5,
                          xmax = 4,
                          ymin = 0.6,
                          ymax = Inf
                        ) +
                        ggplot2::annotation_custom(
                          langosta.grob,
                          xmin = 3.5,
                          xmax = 4,
                          ymin = 0.52,
                          ymax = Inf
                        ) +
                        ggplot2::annotation_custom(
                          pulpo.grob,
                          xmin = 3.5,
                          xmax = 4,
                          ymin = 0.35,
                          ymax = Inf
                        ) +
                        ggplot2::annotation_custom(
                          verdillo.grob,
                          xmin = 3.5,
                          xmax = 4,
                          ymin = 0.05,
                          ymax = Inf
                        )
                      
                      
ggplot2::ggsave(
  "riesgo_pesqueria.png",
  topsis.plot.icons,
  path = here::here("outputs", "figures"),
  dpi = 400,
  width = 8,
  height = 8,
  scale = 1
)
                      

```

PCA

```{r get pca scores}

pca.files <- list.files(path=here::here("outputs","analysis"),pattern = "*_contrib.csv", full.names = TRUE)

pca.res <- lapply(pca.files, get_pca) %>% 
  dplyr::bind_rows()

readr::write_csv(pca.res,here::here("outputs", "analysis","results_pca.csv"))

```

